Texture Mapping 3D Models of Indoor Environments with Noisy Camera Poses

### 2012-06-22 3dimpvt/paper.tex
draft 2.5

%%%%%%%%% TITLE
\title{Texture Mapping Indoor Environments with Imperfect Camera Localization}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying realistic textures to these models is an important
  step in generating accurate visualizations of data gathered by
  modeling systems.  Unfortunately, the localization of cameras in
  such systems often suffer from inaccuracies, resulting in visible
  discontinuties when different images are projected adjacently onto a
  plane for texturing. Previous approaches at minimizing these
  discontinuities do not robustly handle a wide range of camera
  locations and angles, and often suffer from error accumulation when
  stitching together multiple images. We propose two approaches for
  reducing discontinuities during texture mapping, one to robustly
  accomodate images of all orientations, and one to take advantage of
  optimal situations where images have uniform orientation.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}
Three-dimensional modeling of indoor environments has a variety of
applications such as training and simulation for disaster management,
virtual heritage conservation, and mapping of hazardous sites. Manual
construction of these digital models can be time consuming, and as
such, automated 3D site modeling has garnered much interest in recent
years.

The aim of this paper is to present a solution for texture mapping the
3D models generated by indoor modeling systems, with specific
attention given to a human-operated system with higher localization
errors and greater variance in camera locations. The paper is
organized as follows. Section 2 provides an overview of the backpack
modeling system and data processing pipeline from which input data and
examples used throughout this paper originate. Section 3 describes the
general problem of texture mapping and examines two simple mapping
approaches and their problems.  Section 4 demonstrates previous
attempts at localization refinement, and presents our new approach to
image alignment. Section 5 combines our image alignment algorithm with
a simple image selection approach, as well as with a new method for
image selection that takes advantage of optimal images. Section 6
contains results.

\section{Backpack Modeling System}
The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
calculate camera locations within an environment while simulatenously
reconstructing the 3D structure of the environment itself. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

Mounting such devices on a human-carried platform provides unique
advantages over vehicular-based systems in terms of agility and
portability. Unfortunately, human-operated systems also suffer from a
lack of automation and stability, resulting in much higher data
variance and localization error. As a result, common methods for
texture mapping generally produce poor results, as later shown in
section 3. Before discussing how to overcome these challenges, we will
first provide an overview of the backpack modeling system from which
our test data was obtained.

\subsection{Data Acquisition Hardware}
The backpack-mounted modeling system that captured our data contains
five 2D laser range scanners, two cameras, an orientation sensor, and
an IMU. The laser scanners are mounted orthogonally and have a
30-meter range and a 270$^{\circ}$ field of view. The two cameras are
equipped with fisheye lenses, reaching an approximately 180$^{\circ}$
field of view, and are mounted with one facing left and one facing
right. These cameras take images at the rate of 5 Hz. The orientation
sensor provides orientation parameters at a rate of 180 Hz. The IMU
provides highly accurate measurements of all 6 DOF at 200 Hz, and is
used as a ground truth reference for localization processes.

With the backpack actively scanning, the human operator wearing it
takes great care to walk a path such that every wall in the desired
indoor environment is passed and scanned lengthwise at least once.

\subsection{Environment Reconstruction}
Using data gathered by the onboard sensors and multiple localization
and loop-closure algorithms, the backpack is first localized over its
data collection period, and a 3D point cloud of the surrounding
environment is constructed based on the laser scanner readings
relative to the backpack \cite{chen2010indoor}. Approximate normal
vectors for each point in the point cloud are then calculated by
gathering neighboring points within a small radius and processing them
through principal component analysis. These normal vectors allow for
the classification and grouping of adjacent points into structures
such as walls, ceilings, floors, and staircases. A RANSAC algorithm is
then employed to fit polygonal planes to these structured groupings of
points, resulting in a fully planar model
\cite{sanchez2012point}. This model, consisting of multiple 2D
polygonal planes in 3D space, along with the set of images captured by
the backpack's camera, can be considered the input to our texture
mapping problem.

\section{Simple Texture Mapping}

\begin{figure}
  \centering
  \includegraphics[height=2in]{Projection.pdf}
  \caption{Planes are specified in 3D space by four corners $C_1$ to
    $C_4$. Images are related to each plane through the camera
    matrices $P_{1..M}$. }
  \label{fig:projection}
\end{figure}

In all subsequent sections, we will discuss the process of texture
mapping a single plane, as the texturing of each of our planes is
independent and can be completed in parallel.

The geometry of the texture mapping process for a plane is shown in
figure \ref{fig:projection}.  As described in the previous section, we
are provided with a set of $M$ images with which we must texture our
target plane. Each image has a camera matrix $P_i$ for $i=1..M$, which
translates a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. If the 3D world point is not
contained in the image, the 2D point will simply be outside of the
image boundaries. A camera matrix $P_i$ is composed of the camera's
intrinsic parameters, such as focal length and image center, as well
as extrinsic parameters which specify the rotation and translation of
the camera's position in 3D world coordinates at the time that image
$i$ was taken. These extrinsic parameters are determined by the
backpack hardware and localization algorithms mentioned in Section
2. A point $X$ on the plane in 3D space can be related to its
corresponding pixel $x$ in image $i$ through the following equation:

@@

For the sake of simplicity, we treat all planes as rectangles by
generating minimum bounding boxes for them. Since our final textures
will be stored as standard rectangular images anyway, we can simply
leave the area between plane boundary and bounding box untextured, or
crop it out as needed. A plane to be textured is thus defined by a
bounding box with corners $C_i$ in world coordinates and a normal
vector indicating the front facing side of the plane. Our goal is to
texture this plane using images captured by the backpack, while
eliminating any visual discontinuities or seams that would suggest
that the plane's texture was not composed of a single continuous
image.

\subsection{Direct Mapping}

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
generally about 5 pixels across, and picking an image to texture each
tile with. Since we are provided with far more images than necessary
to cover our plane, this process is a simple way of assigning
texturess to locations on our plane. We choose to work with
rectangular units to ensure that borders between any two distinct
images in our final texture will be either horizontal or
vertical. Since most environmental features inside buildings are
horizontal or vertical, any seams in our texture will intersect them
miniimally and be less noticeable.

In order to select an image for texturing tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
and using the projection method above. Furthermore, each candidate
image must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using
standard ray-polygon intersection tests between the camera location,
the center of $t$, and other planes, all in world coordinates.

Once we have a list of candidate images for $t$, we must define a
scoring function in order to compare images and objectively select the
best one. Since camera localization errors compound over distance, we
wish to minimize the distance between cameras used for texturing and
our plane. Additionally, we desire images that project squarely onto
the plane, maximizing the resolution and amount of useful texture
available in their projections.

A scoring function to maximize these two criteria is shown in figure
\ref{fig:scoringFunction}.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{For each tile, we find the image that minimizes $d$ and
    $\alpha$ by maximizing the following equation: \\ $(1/d \cdot (-1
    \cdot C) \cdot N)$}
  \label{fig:scoringFunction}
\end{figure}

@@

\subsection{Mapping with Caching}
Since discontinuities occur where adjacent tiles select different
images that do not match up well, it makes sense to take into account
image choices made by neighboring tiles while selecting the best image
for a tile. By using the same image across tile boundaries, we
eliminate the discontinuity altogether. If this is not possible, using
very similar images will result in less noticeable discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate,
before searching through the entire set. The first subset of images is
the set of images selected by adjacent tiles that have already been
processed. We must first check which images can map to $t$, and then
of those, we make a choice according to the same scoring function in
figure \ref{fig:scoringFunction}. Rather than blindly reusing this
image, we ensure it meets a predefined score threshold to be
considered a good image. If no good image is found, we then check our
second subset of images, which consists of images that were taken near
the images in the first subset, both spatially and temporally. These
images are not the same as the ones used for neighboring tiles, but
they were taken at a similar location and time, suggesting that their
localization and projection will be very similar. Again, if no good
image is found according to a threshold, we then must search the
entire set of candidate images.

@@

The results of this caching approach are shown in figure
\ref{fig:caching}.  Discontinuities have been reduced overall, but the
amount of remaining seams suggests that image selection alone cannot
produce seamless textures. The image projections themselves will have
to be adjusted in order to reliably generate clean textures.

\section{Localization Refinement}
In order to produce photorealistic texture mapping, either camera
matrices need to be refined such that their localization is pixel
accurate, resulting in a perfect mapping, or image stitching
techniques need to be applied to provide this illusion. Before
examining these approaches, we first obtain a set of images to work
with.

Rather than perform camera or image adjustments across many thousands
of images, we choose to work with the more limited set of images
corresponding to those chosen by the direct mapping approach, without
caching. This set of images constitutes a good candidate set for
generating a final seamless texture, since it meets three important
criteria. First, this set of images contains at least one image that
covers each tile on our plane. Thus, unless extreme changes in
localization occur, we can ensure that there will be no holes in our
final texture. Second, since images are all selected according to the
same scoring function in figure \ref{fig:scoringFunction}, we know
that our images are all taken at as much of a head-on angle as
possible and should project onto the plane in similar ways. Third, as
a side result of the scoring function, selected images are only good
candidates for the tiles near their center of projection. Thus, there
should be plenty of overlap between selected images, allowing for some
degree of shifting without resulting in holes, as well as area for
blending between them. With this set of images, we now examine two
approaches towards refining and combining their projections, before
demonstrating our own.

\subsection{Image Mosaicing}

\begin{figure}
  \centering
  \includegraphics[width=3in]{panoMy.jpg}
  \caption{Image mosaicing. }
  \label{fig:mosaic}
\end{figure}

@@

Figure \ref{fig:mosaic} shows the output of the AutoStitch software
package which does homography-based image mosaicing. This plane is
nearly a best-case scenerio with many features spread uniformly across
it. Even so, the mosaicing produces errors that causes straight lines
to appear as waves on the plane. This image was generated after
careful hand tuning as well. Many planes that had fewer features
simply failed outright. This leads to the conclusion that image
mosaicing is not a robust enough solution for reliably texture mapping
our dataset.

\subsection{Image-Based 3D Localization Refinement}

\begin{figure}
  \centering
  \includegraphics[width=3in]{Graph_crop.pdf}
  \caption{Using the graph-based localization refinement algorithm
    from [11] suffers from the problem of compounding errors. }
  \label{fig:graph}
\end{figure}

Another approach is to refine the camera matrices using image
correspondences to guide the process. Each image's camera matrix has 6
degrees of freedom that can be adjusted. Previous work on this problem
attempted to refine camera matrices by solving a non-linear
optimization problem \cite{liu2010indoor}. This process is specific to
the backpack system which generated our dataset, as it must be run
during backpack
localization\cite{liu2010indoor,chen2010indoor}. Unfortunately, this
approach suffers from a similar error propagation problem shown in
Figure \ref{fig:graph}. In our new approach, we also refine the
placement of images using image correspondences. However, we do so in
two dimensions on the plane whereas this previous work did so over all
6 degrees of freedom. Refining in two dimensions on the plane is less
flexible in that it does not address projection errors, however, it
provides significant benefits while avoiding the error propagation
problem.

\subsection{Improved Localization through 2D Matching}
Our approach begins with the projection of all our images onto
separate copies of our plane, such that no projected data is covered
up and lost. This is done in the same way as the approaches in section
3. We then perform Hough transforms to rotate these projections
similarly, and then shift each image's projection such that its
overlapping regions with other projections are as seamless as
possible.

\subsection{Hough Transforms}
Our first step is to use Hough transforms to adjust the rotation of
projected images. Rather than try and match the orientation of
features in each image, we simply apply rotations such that any strong
near-vertical features are made completely vertical. This is effective
for indoor modeling, since strong features in indoor scenes usually
contain vertical lines corresponding to doors, wall panels,
rectangular frames, etc. If features in the environment are not
vertical, or there are few features overall, this step should be
skipped.

@@

\subsection{SIFT Feature Matching}
Our next step is to try and fix misalignment between overlapping
images. We do this by first searching for corresponding points between
all pairs of overlapping images using SIFT feature matching
\cite{lowe1999object}. An illustration of this is given in Figure
\ref{fig:matches}. The SIFT matches allow us to determine
feature-based $x$ and $y$ distances between two images on the plane,
which will allow us to reevaluate where each image should be
projected.

\subsubsection{Robust SIFT Distances using RANSAC}
Since indoor environments often contain repetitive features, such as
floor tiles or doors, we need to ensure that our SIFT-based distances
are reliable. In order to diminish the effect of incorrect matches and
outliers, the RANSAC framework \cite{fischler1981random} is used for a
robust estimate of the $x$ and $y$ distances between the two
images. The RANSAC framework attempts to build a consensus among the
SIFT matches between every pair of images, and tries to generate the
horizontal and vertical distance between them while ignoring the
influence of outliers that would skew the results. The framework
handles the consensus-building machinery, and only requires that two
functions be specified: the fitting function and the distance
function. These functions are called for random subsets of the SIFT
matches until the best set of inliers is found. For this application,
the fitting function simply finds the average distance between
matches. If the matches are exactly correct and the image is frontal
and planar then the distances for various SIFT feature matches should
be the same. Our distance function for a pair of points is the
difference between those points' SIFT match distance and the average
distance computed by the fitting function. We specified a 10 pixel
outlier threshold to the framework. This means that a SIFT match is
labeled as an outlier if its horizontal or vertical distance is not
within 10 pixels of the average distance computed by the fitting
function.

\subsubsection{Refining Image Positions using Least Squares}

There are a total of $M^{2}$ possible pairs of images, though we only
generate distances between images that overlap at SIFT feature
points. Given these distances and the original image location
estimates, we can solve a least squares problem ($\textrm{min}_{\beta}
||\beta X - y||_2^2 $) to estimate the correct location of the images
on the plane. The vector $\beta$ of unknowns represent the correct $x$
and $y$ locations of each image on the plane from $1 \dots M$. The
optimal $x$ and $y$ locations are calculated in the same way, so we
will only consider the $x$ locations here:

@@

\[\beta =
\begin{pmatrix}
  x_1, & x_2, & x_3, & \cdots & x_{M-1}, & x_M
\end{pmatrix}
\]

The matrix $X$ is constructed with one row for each pair of images
with measured distances produced by the SIFT matching stage. A row in
the matrix has a $-1$ and $1$ in the columns corresponding to the two
images in the pair. For example, the matrix below indicates that we
generated a SIFT-based distance between images 1 and 2, images 1 and
3, images 2 and 3, etc.

@@

Finally, the observation vector $y$ is constructed using the
SIFT-based distances generated earlier in the matching stage. The
distances are denoted as $d_1 \dots d_N$ for $N$ SIFT-based
distances. The last element in the observation vector is the location
of the first image determined by its original localization.

\[
y^T =
\begin{pmatrix}
  d_{1,2}, &d_{1,3}, &d_{2,3}, &\hdots &d_{N-2,N-1}, &d_{N-1,N}, &x_1
\end{pmatrix}
\]

The $\beta$ that minimizes $||\beta X - y||_2^2$ results in a set of
image locations on the plane that best honors all the SIFT-based
distance measurements between images. In practice there are often
cases where there is a break in the chain of images, meaning that no
SIFT matches were found between one segment of the plane and
another. In this case we add rows to the $X$ matrix and observations
to the $y$ vector that contain the original $x$ and $y$ distance
estimates generated by the localization algorithm. Another way to do
this is to add rows for all neighboring pairs of images and solve a
weighted least squares problem where the SIFT distances are given a
higher weight i.e. 1, and the distances generated by the localization
algorithm are given a smaller weight i.e. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our image projections should overlap and
match eachother with far greater accuracy.

\section{Improved Texture Mapping with Localization and Blending}
Now that our images have been relocalized for much greater accuracy
relative to eachother, we can revisit the cached mapping approach from
section 3.2. Since we now have higher confidence in the quality of
image boundaries, we can also explore an alternate method that greatly
reduces seams in cases where we have an abundance of optimal images.

\subsection{Localized Mapping with Caching}
In figure \ref{fig:shifted}, we see the same wall from section 3,
texture mapped using the caching method after our localization
refinement process.

@@

It should be evident that our localization adjustments resulted in a
major improvement. A great strength of our tiling approach for image
selection is that it extracts the areas with good camera angles from
each image, and thus can handle images of all types. As demonstrated
on the floor plane in figure \ref{fig:floor_suboptimal}, we can run
this process even with extremely poor image projections, where the
amount of quality texture for each image is limited. Additionally,
since our plane is discretized, we can handle arbitrarily
complex plane geometry as well.

\begin{figure}
  \centering
  \includegraphics[width=3in]{floor_suboptimal.jpg}
  \caption{The tiling method ensures that only the most optimal
    portions of each image are used. This allows it to successfully
    handle all manner of plane and image geometry, as shown in this
    floor texture.}
  \label{fig:floor_suboptimal}
\end{figure}

Given that the backpack system's hardware and the data collection
process strive for optimal wall texturing however, we can aim for a
less conservative approach when texturing walls, and simply fall back
to this more robust approach as needed.

>>>>

%%%%%%%%% TITLE
\title{Texture Mapping 3D Models of Indoor Environments with Noisy
  Camera Poses}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying textures to these models is an important step in
  generating photorealistic visualizations of data gathered by
  modeling systems.  The localization of cameras in such systems often
  suffer from inaccuracies, resulting in visible discontinuties when
  different images are projected adjacently onto a plane for
  texturing. We propose two approaches for reducing discontinuities
  during texture mapping, one to robustly accomodate images of all
  orientations, and one to take advantage of optimal situations where
  images have uniform orientation. The effectiveness of our approaches
  will be demonstrated on two indoor datasets.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}
\label{sec:introduction}
Three-dimensional modeling of indoor environments has a variety of
applications such as training and simulation for disaster management,
virtual heritage conservation, and mapping of hazardous sites. Manual
construction of these digital models can be time consuming, and as
such, automated 3D site modeling has garnered much interest in recent
years.

The aim of this paper is to present a solution for texture mapping the
3D models generated by indoor modeling systems, with specific
attention given to a human-operated system with high camera pose
errors and great variance in camera locations.

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover camera poses within an environment while simulatenously
reconstructing the 3D structure of the environment itself. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

Mounting such devices on a human-carried platform provides unique
advantages over vehicular-based systems in terms of agility and
portability. Unfortunately, human-operated systems also result in much
larger localization error. As a result, common methods for texture
mapping generally produce poor results, as later shown in Sections
\ref{sec:simpleTextureMapping} and
\ref{sec:existingApproaches}. Before discussing how to overcome these
challenges, we first provide an overview of the backpack modeling
system from which our test data has been obtained.

Our backpack-mounted modeling system contains five 2D laser range
scanners, two cameras, and an orientation sensor. The laser scanners
are mounted orthogonally and have a 30-meter range and a 270$^{\circ}$
field of view. The two cameras are equipped with fisheye lenses,
reaching an approximately 180$^{\circ}$ field of view, and are mounted
with one facing left and the other facing right. These cameras take
images at the rate of 5 Hz. The orientation sensor provides
orientation parameters at a rate of 180 Hz.

With the laser scanners active, the human operator wearing the
backpack takes great care to walk a path such that every wall in the
desired indoor environment is traversed and scanned lengthwise at
least once.

Using data gathered by the onboard sensors and multiple localization
and loop-closure algorithms, the backpack is then localized over its
data collection period, and a 3D point cloud of the surrounding
environment is constructed \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}. Approximate normal vectors for each point in the
point cloud are then calculated by gathering neighboring points within
a small radius and processing them through principal component
analysis. These normal vectors allow for the classification and
grouping of adjacent points into structures such as walls, ceilings,
floors, and staircases. A RANSAC algorithm is then employed to fit
polygonal planes to these structured groupings of points, resulting in
a fully planar model \cite{sanchez2012point}. This model, consisting
of multiple 2D polygonal planes in 3D space, along with the set of
images captured by the backpack's cameras and their 3D poses, can be
considered the input to our texture mapping problem

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} describes the general problem of
texture mapping and examines two simple mapping approaches and their
shortcomings.  Section \ref{sec:existingApproaches} demonstrates two
previous attempts at localization refinement, and demonstrates their
inadequacies for our datasets.  Section \ref{sec:proposedApproach}
presents our proposed approach to texture mapping, combining an
improved localization refinement process with two image selection
approaches. Section \ref{sec:resultsAndConclusions} contains results
and conclusions.

\section{Simple Texture Mapping}
\label{sec:simpleTextureMapping}
\begin{figure}
  \centering
  \includegraphics[height=2in]{Projection.pdf}
  \caption{Planes are specified in 3D space by four corners $C_1$ to
    $C_4$. Images are related to each plane through the camera
    matrices $P_{1..M}$. }
  \label{fig:projection}
\end{figure}

In all subsequent sections, we discuss the process of texture mapping
a single plane, as the texturing of each of our planes is independent
and can be completed in parallel.

The geometry of the texture mapping process for a plane is shown in
Figure \ref{fig:projection}.  As described earlier, we are provided
with a set of $M$ images with which we must texture our target
plane. Each image has a camera matrix $P_i$ for $i=1..M$, which
translates a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. If the 3D world point is not
contained in the image, the 2D point will simply be outside of the
image boundaries. A camera matrix $P_i$ is composed of the camera's
intrinsic parameters, such as focal length and image center, as well
as extrinsic parameters which specify the rotation and translation of
the camera's position in 3D world coordinates at the time that image
$i$ is taken. These extrinsic parameters are determined by the
backpack hardware and localization algorithms mentioned earlier. A
point $X$ on the plane in 3D space can be related to its corresponding
pixel $x$ in image $i$ through the following equation:

@@

For the sake of simplicity, we treat all planes as rectangles by
generating minimum bounding boxes for them. Since our final textures
are stored as standard rectangular images anyway, we can simply leave
the area between plane boundary and bounding box untextured, or crop
it out as needed. A plane to be textured is thus defined by a bounding
box with corners $C_i$ in world coordinates and a normal vector
indicating the front facing side of the plane. Our goal is to texture
this plane using images captured by the backpack system, while
eliminating any visual discontinuities or seams that would suggest
that the plane's texture is not composed of a single continuous image.

\subsection{Direct Mapping}
\label{sec:directMapping}

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
generally about 5 pixels across, and choosing an image to texture each
tile with. We choose to work with rectangular units to ensure that
borders between any two distinct images in our final texture are
either horizontal or vertical. Since most environmental features
inside buildings are horizontal or vertical, any seams in our texture
intersect them miniimally and are likely to be less noticeable.

In order to select an image for texturing tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
using the projection method above. Furthermore, each candidate image
must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using standard
ray-polygon intersection tests between the camera location, the center
of $t$, and other planes, all in world coordinates.

Once we have a list of candidate images for $t$, we must define a
scoring function in order to compare images and objectively select the
best one. Since camera localization errors compound over distance, we
wish to minimize the distance between cameras used for texturing and
our plane. Additionally, we desire images that are projected
perpendicularly onto the plane, maximizing the resolution and amount
of useful texture available in their projections.

These two criteria can be met by maximizing the function

\[
\frac{1}{d} \cdot (-1 \cdot C) \cdot N
\]

which minimizes camera angle $\alpha$ and distance $d$ as shown in
Figure \ref{fig:scoringFunction}.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{Images are selected by minimizing camera angle $\alpha$ and
    distance $d$.}
  \label{fig:scoringFunction}
\end{figure}

@@

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select different
images that do not align, it makes sense to take into account image
selections made by neighboring tiles while selecting the best image
for a given tile. By using the same image across tile boundaries, we
eliminate the discontinuity altogether. If a tile is not visible in
images chosen by its neighbors, using similar images is likely to
result in less noticeable discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate,
before searching through the entire set. The first subset of images is
those selected by adjacent tiles that have already been textured. We
must first check which images can map to $t$, and then of those, we
make a choice according to the same scoring function in Figure
\ref{fig:scoringFunction}. Rather than blindly reusing this image, we
ensure it meets a threshold, which we set to $\alpha < 45^\circ$, to
be considered a good image.

If no good image is found, we then check our second subset of images,
which consists of images that were taken near the images in the first
subset, both spatially and temporally. These images are not the same
as the ones used for neighboring tiles, but they were taken at a
similar location and time, suggesting that their localization and
projection are very similar. Again, if no good image is found
according to the same threshold, we then must search the entire set of
candidate images.

@@

The result of this caching approach is shown in figure
\ref{fig:caching}. As compared to \ref{fig:directMapping},
discontinuities have been reduced overall, but the amount of remaining
seams suggests that image selection alone cannot produce seamless
textures. Camera matrices, or the image projections themselves have to
be adjusted in order to reliably generate clean textures.

\section{Existing Approaches to Image-Aligned Texture Mapping}
\label{sec:existingApproaches}
In order to produce seamless texture mapping, either camera matrices
need to be refined such that their localization is pixel accurate,
resulting in a perfect mapping, or image stitching techniques need to
be applied to provide this illusion. Before examining these
approaches, we first obtain a set of images to work with.

Rather than perform camera or image adjustments across the many
thousands of images acquired in a typical data collection, we opt to
work with the more limited set of images corresponding to those chosen
by the direct mapping approach, without caching. This set of images
constitutes a good candidate set for generating a final seamless
texture since it meets three important criteria. First, it contains at least one image that covers each tile on our
plane; this ensures no holes in our final texture. Second, since
images are all selected according to the same scoring function in
Figure \ref{fig:scoringFunction}, they are taken at as much of a head-on angle as possible and should project
onto the plane in similar ways. Third, as a side result of the scoring
function, selected images are only good candidates for the tiles near
their center of projection. Thus, there should be plenty of overlap
between selected images, allowing for some degree of shifting without
resulting in holes, as well as area for blending between them. With
this set of images, we now review two existing approaches towards refining and
combining their projections.

\subsection{Image Mosaicing}
\label{sec:imageMosaicing}
\begin{figure}
  \centering
  \includegraphics[width=3in]{panoMy.jpg}
  \caption{Image mosaicing. }
  \label{fig:mosaic}
\end{figure}

@@

Figure \ref{fig:mosaic} shows the output of the AutoStitch software
package which does homography-based image mosaicing. This plane is
nearly a best-case scenerio with many features spread uniformly across
it. Even so, the mosaicing produces errors that causes straight lines
to appear as waves on the plane. This image was generated after
careful hand tuning as well. Many planes that had fewer features
simply failed outright. Thus, image mosaicing is not a robust enough
solution for reliably texture mapping our dataset.

\subsection{Image-Based 3D Localization Refinement}
\label{sec:imageBased3DRefinement}
\begin{figure}
  \centering
  \includegraphics[width=3in]{Graph_crop.pdf}
  \caption{The graph-based localization refinement algorithm from [11]
    suffers from the problem of compounding errors. }
  \label{fig:graph}
\end{figure}

Another approach is to refine the camera matrices using image
correspondences to guide the process. Each image's camera matrix has 6
degrees of freedom that can be adjusted. Previous work on this problem
attempted to refine camera matrices by solving a non-linear
optimization problem \cite{liu2010indoor}. This process is specific to
the backpack system which generated our dataset, as it must be run
during backpack
localization\cite{liu2010indoor,chen2010indoor}. Unfortunately, this
approach suffers from a similar error propagation problem shown in
Figure \ref{fig:graph}.

\section{Proposed Method for Seamless Texture Mapping}
\label{sec:proposedApproach}
Our proposed approach towards texture mapping is a two-step
process. First, with the same input set of images as described in
Section \ref{sec:existingApproaches}, we perform image rotation and
shifting in order to maximize SIFT feature matches between images. We
then project these images onto our plane, applying textures either
with the tile caching method from Section
\ref{sec:mappingWithCaching}, or with the more specialized method
described in Section \ref{sec:seamMinimization}.

\subsection{Image Projection and Rotation}
\label{sec:projectionAndRotation}
Our localization approach begins with the projection of all our images
onto separate copies of our plane, such that no projected data is
covered up and lost. This is done in the same way as the approaches in
Section \ref{sec:simpleTextureMapping}. We then perform rotations on
these projections, as adjacent images with different orientations will
result in strong discontinuities.

These rotations are accomplished by using Hough transforms, which
detect the presence and orientation of linear features in our
images. Rather than match the orientation of such features in each
image, we simply apply rotations such that the strongest near-vertical
features are made completely vertical. This is effective for indoor
models, since strong features in indoor scenes usually consist of
parallel vertical lines corresponding to doors, wall panels,
rectangular frames, etc. If features in the environment are not
vertical, or are not parallel to eachother, this step is skipped.

@@

\subsection{Robust SIFT Feature Matching using RANSAC}
\label{sec:robustSIFTFeatureMatching}
Our next step is to fix misalignment between overlapping images. We do
this by first searching for corresponding points between all pairs of
overlapping images using SIFT feature matching
\cite{lowe1999object}. An illustration of this is given in Figure
\ref{fig:matches}. The SIFT matches allow us to determine $dx$ and
$dy$ distances between each pair of features for two images on the
plane, determining where they should be projected relative to
eachother.

Since indoor environments often contain repetitive features such as
floor tiles or doors, we need to ensure that our SIFT-based distances
are reliable. In order to mitigate the effect of incorrect matches and
outliers, the RANSAC framework \cite{fischler1981random} is used for a
robust estimate of the optimal $dx$ and $dy$ distances between two
images. The RANSAC framework handles the consensus-building machinery,
and requires a fitting function and a distance function. For this
application, the fitting function simply finds the average distance
between matches in a pair of images. Our distance function for a pair
of points is the difference between those points' SIFT match distance
and the average distance computed by the fitting function; we use a 10
pixel outlier threshold. This means that a SIFT match is labeled as an
outlier if its horizontal or vertical distance is not within 10 pixels
of the average distance computed by the fitting function.

\subsubsection{Refining Image Positions using Least Squares}
\label{sec:refiningImagePositions}
There are a total of $M^{2}$ possible pairs of images, though we only
generate distances between images that overlap at SIFT feature
points. Given these distances and the original image location
estimates, we can solve a least squares problem ($\textrm{min}_{\vec{\beta}}
||X \vec{\beta} - \vec{\gamma}||_2^2 $) to estimate the correct location of the images
on the plane. The $M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
and $y$ locations of each image on the plane from $1 \dots M$. The
optimal $x$ and $y$ locations are obtained in the same way, so we
only consider the $x$ locations here:

@@

\[\vec{\beta} =
\begin{pmatrix}
  x_1, & x_2, & x_3, & \cdots & x_{M-1}, & x_M
\end{pmatrix}
\]

The $N$ by $(M+1)$ dimensional matrix $X$ is constructed with one row for each pair of images
with measured distances produced by the SIFT matching stage. A row in
the matrix has a $-1$ and $1$ in the columns corresponding to the two
images in the pair. For example, the matrix below indicates that we
generated a SIFT-based distance between images 1 and 2, images 1 and
3, images 2 and 3, etc.

@@

Finally, the $N$-dimensional observation vector $\vec{\gamma}$ is constructed using the
SIFT-based distances generated earlier in the matching stage. The
distances are denoted as $d_1 \dots d_N$ for $N$ SIFT-based
distances. The last element in the observation vector is the location
of the first image determined by its original noisy localization, from \cite{chen2010indoor, liu2010indoor}:

\[
\vec{y}^T =
\begin{pmatrix}
  d_{1,2}, &d_{1,3}, &d_{2,3}, &\hdots &d_{N-2,N-1}, &d_{N-1,N}, &x_1
\end{pmatrix}
\]

The $\vec{\beta}$ that minimizes $||X \vec{\beta} - \vec{\gamma}||_2^2$ results in a set of
image locations on the plane that best honors all the SIFT-based
distance measurements between images. In practice there are often
cases where there is a break in the chain of images, meaning that no
SIFT matches were found between one segment of the plane and
another. In this case we add rows to the $X$ matrix and observations
to the $\vec{\gamma}$ vector that contain the original noisy $x$ and $y$ distance
estimates generated by the localization algorithm \cite{chen2010indoor, liu2010indoor}. Another way to do
this is to add rows for all neighboring pairs of images and solve a
weighted least squares problem where the SIFT distances are given a
higher weight i.e. 1, and the noisy distances generated by the localization
algorithm \cite{chen2010indoor, liu2010indoor} are given a smaller weight i.e. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and
match each other with far greater accuracy. Applying the tile caching method from Section \ref{sec:mappingWithCaching} on these re-localized images, results in the significant improvements shown in Figure \ref{fig:shifted}.

@@

\subsection{Texture Mapping with Seam Minimization}
\label{sec:seamMinimization}

As mentioned earlier, our backpack system has 2 cameras with
$180^\circ$ fisheye lenses facing to the right and left. Since the
backpack operator only takes care to fully scan each wall and not
necessarily the entirety of each ceiling and floor, cameras at higher
angles with respect to plane normal vectors must be used for texturing
large areas of floor and ceiling planes. These high camera angles
translate into images that span extremely large areas once projected
onto their planes. Such image projections have good scores (from
Figure \ref{fig:scoringFunction}) for certain areas on the plane, but
much worse scores for other areas. This can be seen in Figure
\ref{fig:highCameraAngle}. The tiling approach used in Section
\ref{sec:simpleTextureMapping} was thus a good choice for texturing,
as it allowed us to only use the parts of image projections that were
good choices for their respective plane locations.

For wall planes however, we have images all taken from close distances
and more head-on angles, and thus much smaller near-rectangular
projections to work with. As a result, there is less deviation over
the score of each tile within an image, as well as over the average
scores of all images. This means that the scoring criteria from Figure
\ref{fig:scoringFunction} is less relevant to walls, which enjoy an
abundance of head-on images.

Thus, for planes with optimal images, rather than selecting the set of
best images, since all images are near in quality, we instead select
the best set of images, such that the selection together results in
the cleanest final texture. We will accomplish this by using entire
images where possible, and defining a cost function to minimize the
visibility of seams in our final texture.
